Nonlinear Models
========================================================
Here we explore the use of nonlinear models using some tools in R

```{r}
require(ISLR)
attach(Wage)
```

Polynomials
------------

First we will use polynomials, and focus on a single predictor age:

```{r}
fit=lm(wage~poly(age,4),data=Wage)
summary(fit)
coef(fit)
```

The `poly()` function generates a basis of *orthogonal polynomials*.
Lets make a plot of the fitted function, along with the standard errors of the fit.

```{r fig.width=7, fig.height=6}
agelims=range(age)
age.grid=seq(from=agelims[1],to=agelims[2])
preds=predict(fit,newdata=list(age=age.grid),se=TRUE)
se.bands=cbind(preds$fit+2*preds$se,preds$fit-2*preds$se)
plot(age,wage,col="darkgrey")
lines(age.grid,preds$fit,lwd=2,col="blue")
matlines(age.grid,se.bands,col="blue",lty=2)
```

Since f(x0) is a linear function of the betta, can get a simple expression for *pointwise-variances* Var[f(x0)] at any value x0. In the figure we have computed the fit and pointwise standard errors on a grid of values for x0. We show
f(x0) = +- 2*se[f(x0)]

There are other more direct ways of doing this in R. For example

```{r}
fita=lm(wage~age+I(age^2)+I(age^3)+I(age^4),data=Wage)
summary(fita)
coef(fita)
```

Here `I()` is a *wrapper* function; we need it because *age^2* means something to the formula language,
while `I(age^2)` is protected.
The coefficients are different to those we got before! However, the fits are the same:

```{r}
plot(fitted(fit),fitted(fita))
```

By using orthogonal polynomials in this simple way, it turns out that we can separately test
for each coefficient. So if we look at the summary again, we can see that the linear, quadratic
and cubic terms are significant, but not the quartic.

```{r}
summary(fit)
```

This only works with linear regression, and if there is a single predictor. In general we would use **anova()** as this next example demonstrates.

In performing a polynomial regression we must decide on the degree of the polynomial to use.

 We use the **anova()** function, which performs an **anova()** analysis of variance (ANOVA, using an F-test) in order to test the null hypothesis that a model **M1** is sufficient to explain the data against the alternative hypothesis that a more complex model **M2** is required.
 In order to use the anova() function, **M1** and **M2** must be nested models: the predictors in **M1** must be a subset of the predictors in M2.
 In this case, we fit five different models and sequentially compare the simpler model to the more complex model.
```{r}
fita=lm(wage~education,data=Wage)
fitb=lm(wage~education+age,data=Wage)
fitc=lm(wage~education+poly(age,2),data=Wage)
fitd=lm(wage~education+poly(age,3),data=Wage)
anova(fita,fitb,fitc,fitd)

```
 The p-value comparing the linear **Model 1** to the quadratic **Model 2** is essentially zero (<10-15), indicating that a linear fit is not sufficient.
 Similarly the p-value comparing the quadratic **Model 2** to the cubic **Model 3**
is very low (0.0017), so the quadratic fit is also insufficient.
 The p-value comparing the cubic and degree-4 polynomials, **Model 3** and **Model 4**, is approximately 5% while the degree-5 polynomial **Model 5** seems unnecessary
because its p-value is 0.37.
 Hence, either a cubic or a quartic polynomial appear to provide a reasonable fit to the data, but lower- or higher-order models are not justified.


### Polynomial logistic regression

Now we fit a logistic regression model to a binary response variable,
constructed from `wage`. We code the big earners (`>250K`) as 1, else 0.

```{r}
fit=glm(I(wage>250) ~ poly(age,3), data=Wage, family=binomial)
summary(fit)
preds=predict(fit,list(age=age.grid),se=T)
se.bands=preds$fit + cbind(fit=0,lower=-2*preds$se,upper=2*preds$se)
se.bands[1:5,]
```

We have done the computations on the logit scale. To transform we need to apply the inverse logit
mapping
$$p=\frac{e^\eta}{1+e^\eta}.$$
(Here we have used the ability of MarkDown to interpret TeX expressions.)
We can do this simultaneously for all three columns of `se.bands`:

```{r}
prob.bands=exp(se.bands)/(1+exp(se.bands))
matplot(age.grid,prob.bands,col="blue",lwd=c(2,1,1),lty=c(1,2,2),type="l",ylim=c(0,.1))
points(jitter(age),I(wage>250)/10,pch="|",cex=.5)
```
We have drawn the age values corresponding to the observations with wage values above 250 as gray marks on the top of the plot, and those with wage values below 250 are shown as gray marks on the bottom of the plot.

```{r}
table(cut (age ,4))
fit=lm(wage~cut (age ,4) ,data=Wage)
coef(summary (fit))
```
Here cut() automatically picked the cutpoints at 33.5, 49, and 64.5 years of age. We could also have specified our own cutpoints directly using the breaks option. The function cut() returns an ordered categorical variable; the lm() function then creates a set of dummy variables for use in the regression.

Splines
-------
 Splines are more flexible than polynomials, but the idea is rather similar.
Here we will explore cubic splines.
 A cubic spline with knots is a piecewise cubic polynomial with continuous derivatives up to order 2 at each knot.
 The **bs()** function generates the entire matrix of basis functions for splines with the specified set of knots. By default, cubic splines are produced.
```{r}
require(splines)
fit=lm(wage~bs(age,knots=c(25,40,60)),data=Wage)
plot(age,wage,col="darkgrey")
lines(age.grid,predict(fit,list(age=age.grid)),col="darkgreen",lwd=2)
abline(v=c(25,40,60),lty=2,col="darkgreen")

pred=predict (fit ,newdata =list(age =age.grid),se=T)
lines(age.grid ,pred$fit +2* pred$se ,lty ="dashed")
lines(age.grid ,pred$fit -2* pred$se ,lty ="dashed")
```
 Here we have prespecified knots at ages 25, 40, and 60. This produces a spline with six basis functions. (Recall that a cubic spline with three knots has seven degrees of freedom; these degrees of freedom are used up by an intercept, plus six basis functions.)

 In order to instead fit a natural spline, we use the **ns()** function. Here we fit a natural spline with four degrees of freedom.

```{r}
plot(age, wage, col = "darkgrey")
fit2=lm(wage~ns(age ,df =4) ,data=Wage)
pred2=predict (fit2 ,newdata =list(age=age.grid),se=T)
lines(age.grid , pred2$fit ,col ="red",lwd =2)

```

 The smoothing splines does not require knot selection, but it does have a smoothing parameter,
which can conveniently be specified via the effective degrees of freedom or `df`.

```{r}
fit=smooth.spline(age,wage,df=16)
lines(fit,col="red",lwd=2)
fit$df
```

Or we can use LOO cross-validation to select the smoothing parameter for us automatically:

```{r}
fit=smooth.spline(age,wage,cv=TRUE)
lines(fit,col="purple",lwd=2)
fit$df
fit
```
 Notice that in the first call to **smooth.spline()**, we specified df=16. The function then determines which value of $lambda$ leads to 16 degrees of freedom.
 In the second call to smooth.spline(), we select the smoothness level by crossvalidation; this results in a value of $lambda$ that yields 6.8 degrees of freedom.

Local Regression
---------------------------

```{r}
plot(age ,wage ,xlim=agelims ,cex =.5, col =" darkgrey ")
title (" Local Regression ")
fit=loess (wage~age ,span =.2, data=Wage)
fit2=loess(wage~age ,span =.5, data=Wage)
lines(age.grid ,predict (fit ,data.frame(age=age.grid)),
col ="red ",lwd =2)
lines(age.grid ,predict (fit2 ,data.frame(age=age.grid)),
col =" blue",lwd =2)
legend ("topright",legend =c("Span =0.2" ,"Span =0.5") ,
col=c("red","blue"),lty =1, lwd =2, cex =.8)
```

 Here we have performed local linear regression using spans of 0.2 and 0.5: that is, each neighborhood consists of 20% or 50% of the observations.
 The larger the span, the smoother the fit. The locfit library can also be used
for fitting local regression models in R.

Generalized Additive Models
---------------------------

gam1=lm(wage~ns(year ,4)+ns(age ,5) +education ,data=Wage)

 So far we have focused on fitting models with mostly single nonlinear terms.
The `gam` package makes it easier to work with multiple nonlinear terms. In addition
it knows how to plot these functions and their standard errors.

 The **s()** function, which is part of the gam library, is used to indicate that we would like to use a ****smoothing spline****. We specify that the function of year should have 4 degrees of freedom, and that the function of age will have 5 degrees of freedom.
 
```{r fig.width=10, fig.height=5}
require(gam)
gam1=gam(wage~s(age,df=4)+s(year,df=4)+education,data=Wage)
par(mfrow=c(1,3))
plot(gam1,se=T)
gam2=gam(I(wage>250)~s(age,df=4)+s(year,df=4)+education,data=Wage,family=binomial)
plot(gam2)

plot(gam1, se=TRUE ,col ="blue")
plot.gam(gam1 , se=TRUE , col ="red")
```

 In these plots, the function of year looks rather linear. We can perform a series of **ANOVA** tests in order to determine which of these three models is best: a GAM that excludes year (M1), a GAM that uses a linear function of year (M2), or a GAM that uses a spline function of year (M3).

```{r}
gam.m1=gam(wage~s(age ,5) +education ,data=Wage)
gam.m2=gam(wage~year+s(age ,5)+education ,data=Wage)
gam.m3=gam(wage~s(year ,4)+s(age ,5)+education ,data=Wage)
anova(gam.m1 ,gam.m2 ,gam.m3,test="F")
```
 We find that there is compelling evidence that a GAM with a linear function of year is better than a GAM that does not include year at all (p-value=0.00014). However, there is no evidence that a non-linear function
of year is needed (p-value=0.349). In other words, based on the results of this ANOVA, M2 is preferred.

 We can also use local regression fits as building blocks in a GAM, using the lo() function.

```{r}
gam.lo=gam(wage~s(year ,df=4)+lo(age ,span =0.7)+education ,data=Wage)
plot.gam(gam.lo , se=TRUE , col ="green ")
```

Here we have used local regression for the age term, with a span of 0.7.
 We can also use the lo() function to create interactions before calling the gam() function.
For example Lets see if we need a nonlinear terms for year:

```{r}
gam.lo.i=gam (wage~lo(year ,age ,span =0.5) +education , data=Wage)
library (akima)
plot(gam.lo.i)
```

In order to fit a logistic regression GAM, we once again use the I() function in constructing the binary response variable, and set family=binomial.

```{r}
gam.lr=gam(I(wage >250)~year+s(age ,df =5)+education ,family =binomial ,data=Wage)
par(mfrow =c(1,3))
plot(gam.lr,se=T,col ="green")
```
It is easy to see that there are no high earners in the <HS category:
```{r}
table(education ,I(wage >250) )
```

Hence, we fit a logistic regression GAM using all but this category. This
provides more sensible results.
```{r}
gam.lr.s=gam (I(wage >250)~year+s(age ,df=5)+education ,family =
binomial ,data=Wage ,subset =( education !="1. < HS Grad"))
plot(gam.lr.s,se=T,col ="green")
```


```{r}
gam2a=gam(I(wage>250)~s(age,df=4)+year+education,data=Wage,family=binomial)
anova(gam2a,gam2,test="Chisq")
```

One nice feature of the `gam` package is that it knows how to plot the functions nicely,
even for models fit by `lm` and `glm`.

```{r fig.width=10, fig.height=5}
par(mfrow=c(1,3))
lm1=lm(wage~ns(age,df=4)+ns(year,df=4)+education,data=Wage)
plot.gam(lm1,se=T)
```